### 2015-12-23 conclusions.tex
Improve wording
Improve wording and fix typos

We have shown that one of the scale-free \textsc{FTRL}-based the algorithms has
regret $O \left(f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} + \max_{t=1,2,\dots,T}
\|\ell_t\|_* \sqrt{T} \right)$ with respect to any competitor $u \in K$, where
$f$ is any non-negative $1$-strongly convex function defined on the decision
set.  The latter result makes sense even when the decision set $K$ is
unbounded. Notice that with regularizer $f(w) = \frac{1}{2}\|w\|_2^2$ the
regret depends on norm of the competitor $\|u\|_2$ is quadratically. There
exist non-scale-free algorithms \cite{ McMahan-Streeter-2012,
McMahan-Abernethy-2013, Orabona-2013, McMahan-Orabona-2014, Orabona-2014} that
have only $O(\|u\|_2 \sqrt{\log \|u_2\|_2})$ or $O(\|u\|_2 \log \|u\|_2)$
dependency.  These algorithms, however, assume an a priori bound on the norm of
the loss vectors.  This leads to another open problem:
%
\begin{quotation}
\noindent
\emph{Design a scale-free algorithm that has $O(\|u\|_2 \sqrt{\log \|u\|_2})$
dependency on the norm $\|u\|_2$ of the competitor.}
\end{quotation}

>>>>

We have shown that one of the scale-free \textsc{FTRL}-based the algorithms has
regret $O \left(f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} + \max_{t=1,2,\dots,T}
\|\ell_t\|_* \sqrt{T} \right)$ with respect to any competitor $u \in K$, where
$f$ is any non-negative $1$-strongly convex function defined on the decision
set.  The latter result makes sense even when the decision set $K$ is
unbounded. Notice that with regularizer $f(u) = \frac{1}{2}\|u\|_2^2$ the
regret depends on norm of the competitor $\|u\|_2$ is quadratically. There
exist non-scale-free algorithms \cite{McMahan-Streeter-2012,
McMahan-Abernethy-2013, Orabona-2013, McMahan-Orabona-2014, Orabona-2014} that
have only $O(\|u\|_2 \sqrt{\log \|u_2\|_2})$ or $O(\|u\|_2 \log \|u\|_2)$
dependency.  These algorithms assume an a priori bound on the norm of the loss
vectors.  This leads to another open problem:
%
\begin{quotation}
\noindent
\emph{Is there a scale-free algorithm for the decision set $\R^d$, regret of
which has $O(\|u\|_2 \sqrt{\log \|u\|_2})$ dependency on the competitor $u$?}
\end{quotation}

### 2015-12-23 introduction.tex
Minor fixes
Improve wording and fix typos

For this reason, most of the existing OLO algorithms receive as an input---or
explicitly assume---an upper bound $B$ on the norm of the loss vectors.  The
input $B$ is often disguised as the learning rate, the regularization
parameter, or the parameter of strong convexity of the regularizer. Examples of
such algorithms include the \textsc{Hedge}~\cite{Freund-Schapire-1997} or
online projected gradient descent with fixed learning rate.  However, these
algorithms have two obvious drawbacks.

First, they do not come with any regret guarantee for sequences of loss vectors
with norms exceeding $B$. Second, on sequences where the norm of loss vectors
is bounded by $b \ll B$, these algorithms fail to have an optimal regret
guarantee that depends on $b$ rather than on $B$.

@@

There is a clear practical need to design algorithms that adapt automatically
to norms of the loss vectors.  A natural, yet overlooked, design method to
achieve this type of adaptivity is by insisting to have a \textbf{scale-free}
algorithm.  That is, the sequence of decisions of the algorithm does not change
if the sequence of loss vectors is multiplied by a positive constant.  The most
important property of scale-free algorithms is that both their loss and their
regret scale linearly with the maximum norm of the loss vector appearing in the
sequence.

@@

Both \textsc{FTRL} and \textsc{MD} are parametrized by a function called a
\emph{regularizer}. The function is assumed to be strongly convex with respect
to a given norm. Based on different regularizers different algorithms with
different properties can be instantiated. A summary of algorithms for OLO is
presented in Table~\ref{table:results}.  All of them, except for
\textsc{Hedge}, are instances of \textsc{FTRL} or \textsc{MD}.

Scale-free versions of \textsc{MD} include \textsc{AdaGrad
MD}~\cite{Duchi-Hazan-Singer-2011} and~\textsc{Optimistic
MD}~\cite{Rakhlin-Sridharan-2013}. However, these algorithms have non-trivial
regret bounds only when the Bregman divergence associated with the regularizer
is bounded, which happens only when the decision set is bounded as well. As we
show in Section~\ref{subsection:mirror-descent-lower-bound}, this is
unavoidable, since \textsc{AdaGrad MD} and similar algorithms based on
\textsc{MD} can have $\Omega(T)$ regret if Bregman divergence is not bounded.

>>>>

For this reason, most of the existing OLO algorithms receive as an input---or
explicitly assume---an upper bound $B$ on the norm of the loss vectors.  The
input $B$ is often disguised as the learning rate, the regularization
parameter, or the parameter of strong convexity of the regularizer. Examples of
such algorithms include \textsc{Hedge}~\cite{Freund-Schapire-1997} or online
projected gradient descent with fixed learning rate.  However, these algorithms
have two obvious drawbacks.

First, they do not come with any regret guarantee for sequences of loss vectors
with norms exceeding $B$. Second, on sequences of loss vectors with norms
bounded by $b \ll B$, these algorithms fail to have an optimal regret guarantee
that depends on $b$ rather than on $B$.

@@

There is a clear practical need to design algorithms that adapt automatically
to the norms of the loss vectors.  A natural, yet overlooked, design method to
achieve this type of adaptivity is by insisting to have a \textbf{scale-free}
algorithm.  That is, the sequence of decisions of the algorithm does not change
if the sequence of loss vectors is multiplied by a positive constant.  The most
important property of scale-free algorithms is that both their loss and their
regret scale linearly with the maximum norm of the loss vector appearing in the
sequence.

@@

Both \textsc{FTRL} and \textsc{MD} are parametrized by a function called a
\emph{regularizer}. The function is assumed to be strongly convex with respect
to a given norm. Based on different regularizers different algorithms with
different properties can be instantiated. A summary of algorithms for OLO is
presented in Table~\ref{table:results}.  All of them are instances of
\textsc{FTRL} or \textsc{MD}.

Scale-free versions of \textsc{MD} include \textsc{AdaGrad
MD}~\cite{Duchi-Hazan-Singer-2011} and~\textsc{Optimistic
MD}~\cite{Rakhlin-Sridharan-2013}. However, these algorithms have non-trivial
regret bounds only when the Bregman divergence associated with the regularizer
is bounded. In particular, since a bound on the Bregman divergence implies that
decision set is bounded, the regret bounds are vacuous for unbounded sets.  In
fact, as we show in Section~\ref{subsection:mirror-descent-lower-bound}
\textsc{AdaGrad MD} and similar algorithms based on \textsc{MD} have
$\Omega(T)$ regret if the Bregman divergence is not bounded.

### 2015-12-23 mirror-descent.tex
Minor fixes
Improve wording and fix typos

Several existing algorithms are special cases of \textsc{MD}.
For example, \textsc{Exponentially-Weighted Average
Forecaster}~\cite[Chapter~2]{Cesa-Bianchi-Lugosi-2006} with a constant learning
rate $\eta$ for learning with expert advice is an instance of
\textsc{MD} for the probability simplex $K = \{ w \in \R^d ~:~ w_j
\ge 0, \sum_{j=1}^d w_j = 1 \}$ with negative entropy regularizer $R_t(w) =
\frac{1}{\eta} \sum_{j=1}^d w_j \ln w_j$ where $R_t:K \to \R$. Note that this
algorithm is \emph{not} scale-free.

@@

The regret bound for \textsc{Scale-Free Mirror Descent} in the
Corollary~\ref{corollary:regret-scale-free-mirror-descent} depends on
$\sup_{u,v \in K} \Breg_f(u,v)$. In contrast, the regret bounds for
\textsc{AdaFTRL} and \textsc{SOLO FTRL} in
Corollaries~\ref{corollary:ada-ftrl-regret-bound}~and~\ref{corollary:regret-solo-ftrl-bounded-set}
depend on $\sup_{u \in K} f(u)$.  Similarly, the regret bound in
Theorem~\ref{theorem:regret-scale-free-mirror-descent} for \textsc{Scale-Free
MD} depends on $\sup_{v \in K} \Breg_R(u,v)$ and the regret bounds in
Theorems~\ref{theorem:ada-ftrl-regret-bound}~and~\ref{theorem:regret-solo-ftrl}
for \textsc{AdaFTRL} and \textsc{SOLO FTRL} depend on $R(u)$. It is not hard to
show that
\begin{equation}
\label{equation:regularizer-vs-divergence}
\forall u \in K \qquad R(u) \le \sup_{v \in K} \Breg_R(u,v)
\end{equation}
provided that at the minimizer $v^* = \argmin_{v \in K} R(v)$ both $R(v^*)$ and
$\grad R(v^*)$ are zero. Indeed, in that case, $R(u) = \Breg_R(u,v^*) \le
\sup_{v \in K} \Breg_R(u,v)$.

@@

Thus, \eqref{equation:regularizer-vs-divergence} implies that ignoring constant
factors, the regret bound for \textsc{Scale-Free MD} is inferior to the regret
bounds for \textsc{AdaFTRL} and \textsc{SOLO FTRL}.  In fact, it is not hard to
come up with examples where $R(u)$ is finite whereas $\sup_{v \in K}
\Breg_R(u,v)$ is infinite. We mention two such examples. The first example is
$R(w) = \frac{1}{2}\|w\|_2^2$ defined on the whole space $V$, where for any $u
\in V$, $R(u)$ is a finite value but $\sup_{v \in K} \Breg_R(u,v) = \sup_{v \in
V} \frac{1}{2}\|u-v\|_2^2 = +\infty$. The second example is the negative
entropy regularizer $R(w) = \ln d + \sum_{j=1}^d w_j \ln w_j$ defined on the
$d$-dimensional probability simplex $K = \{ w \in \R^d ~:~ w_j \ge 0,
\sum_{j=1}^d w_j = 1 \}$, where for any $u \in K$, $R(u)$ is finite and in fact
lies in the interval $[-\ln d, 0]$ but $\sup_{v \in K} \Breg_R(u,v) = \sup_{v
\in K} \sum_{j=1}^d u_j \ln(u_j/v_j) = + \infty$. We return to these examples
in the following subsection.

@@

The bounds in Theorem~\ref{theorem:regret-scale-free-mirror-descent} and
Corollary~\ref{corollary:regret-scale-free-mirror-descent} are vacuous when
$\Breg_R(u,v)$ is not bounded. One might wonder if the assumption that
$\Breg_R(u,v)$ is bounded is necessary in order for \textsc{Scale-Free MD} to
have a ``small'' regret. We show necessity of this assumption on two
counter-examples.  In these counter-examples, we consider strongly convex
regularizers $R$ such that $\Breg_R(u,v)$ is not bounded and we construct
sequences of loss vectors $\ell_1, \ell_2, \dots, \ell_T$ such that
$\|\ell_1\|_* = \|\ell_2\|_* = \dots = \|\ell_T\|_* = 1$ and \textsc{Scale-Free
MD} has regret $\Omega(T)$ or worse.

The first counter-example is stated as
Theorem~\ref{theorem:first-counter-example} below; proof is
in~\ref{section:mirror-descent-proofs}. The decision set is the whole space
$K=V$ and the regularizer is $R(w) = \frac{1}{2}\|w\|_2^2$. Note that $R(w)$ is
$1$-strongly convex with respect to $\|\cdot\|_2$ and the dual norm of
$\|\cdot\|_2$ is $\|\cdot\|_2$. The corresponding Bregman divergence is
$\Breg_R(u,v) = \frac{1}{2}\|u-v\|_2^2$. The counter-example constructs
sequence of unit-norm loss vectors in the one-dimensional subspace spanned by
the first vector of the standard orthnormal basis.  On such sequence,
Zinkevich's \textsc{GIGA} and both versions of \textsc{AdaGrad MD} have
predictions identical to \textsc{Scale-Free MD}. Hence the lower bound applies
to all four algorithms.

@@

\begin{theorem}[Second Counter-Example]
\label{theorem:second-counter-example}
Let $d \ge 2$, let $V = \R^d$, and let $K = \{ w \in V ~:~ w_j \ge 0,
\sum_{j=1}^d w_j = 1 \}$ be the $d$-dimensional probability simplex.  For any
$T \ge 120$, there exists a sequence of loss vectors $\ell_1, \ell_2, \dots,
\ell_T \in V^*$ such that $\|\ell_1\|_\infty = \|\ell_2\|_\infty = \dots =
\|\ell_T\|_\infty = 1$ and \textsc{Scale-Free Mirror Descent} with regularizer
$R(w) = \sum_{j=1}^d w_j \ln w_j$ satisfies
$$
\Regret_T \ge \frac{T}{6} \; .
$$
\end{theorem}

>>>>

Several existing algorithms are special cases of \textsc{MD}.  For example,
\textsc{Hedge} is an instance of \textsc{MD} for the probability simplex $K =
\{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$ with negative entropy
regularizer $R_t(w) = \frac{1}{\eta} \sum_{j=1}^d w_j \ln w_j$ where $\eta > 0$
is a (constant) learning rate. Note that this algorithm is \emph{not}
scale-free.

@@

The regret bound for \textsc{Scale-Free MD} in the
Corollary~\ref{corollary:regret-scale-free-mirror-descent} depends on
$\sup_{u,v \in K} \Breg_f(u,v)$. In contrast, the regret bounds for
\textsc{AdaFTRL} and \textsc{SOLO FTRL} in
Corollaries~\ref{corollary:ada-ftrl-regret-bound}~and~\ref{corollary:regret-solo-ftrl-bounded-set}
depend on $\sup_{u \in K} f(u)$.  Similarly, the regret bound in
Theorem~\ref{theorem:regret-scale-free-mirror-descent} for \textsc{Scale-Free
MD} depends on $\sup_{v \in K} \Breg_R(u,v)$ and the regret bounds in
Theorems~\ref{theorem:ada-ftrl-regret-bound}~and~\ref{theorem:regret-solo-ftrl}
for \textsc{AdaFTRL} and \textsc{SOLO FTRL} depend on $R(u)$. It is not hard to
show that
\begin{equation}
\label{equation:regularizer-vs-divergence}
\forall u \in K \qquad R(u) \le \sup_{v \in K} \Breg_R(u,v)
\end{equation}
provided that at the minimizer $v^* = \argmin_{v \in K} R(v)$ both $R(v^*)$ and
$\grad R(v^*)$ are zero. Indeed, in that case, $R(u) = \Breg_R(u,v^*) \le
\sup_{v \in K} \Breg_R(u,v)$.

@@

Thus, inequality \eqref{equation:regularizer-vs-divergence} implies
that---ignoring constant factors---the regret bound for \textsc{Scale-Free MD}
is inferior to the regret bounds for \textsc{AdaFTRL} and \textsc{SOLO FTRL}.
In fact, it is not hard to come up with examples where $R(u)$ is finite whereas
$\sup_{v \in K} \Breg_R(u,v)$ is infinite. We mention two such examples. The
first example is $R(w) = \frac{1}{2}\|w\|_2^2$ defined on the whole space $V$,
where for any $u \in V$, $R(u)$ is a finite value but $\sup_{v \in K}
\Breg_R(u,v) = \sup_{v \in V} \frac{1}{2}\|u-v\|_2^2 = +\infty$. The second
example is the shifted negative entropy regularizer $R(w) = \ln(d) +
\sum_{j=1}^d w_j \ln w_j$ defined on the $d$-dimensional probability simplex $K
= \{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$, where for any $u \in
K$, $R(u)$ is finite and in fact lies in the interval $[0, \ln d]$ but
$\sup_{v \in K} \Breg_R(u,v) = \sup_{v \in K} \sum_{j=1}^d u_j \ln(u_j/v_j) = +
\infty$. We revisit these examples in the following subsection.

@@

The bounds in Theorem~\ref{theorem:regret-scale-free-mirror-descent} and
Corollary~\ref{corollary:regret-scale-free-mirror-descent} are vacuous when
$\Breg_R(u,v)$ is not bounded. One might wonder if the assumption that
$\Breg_R(u,v)$ is bounded is necessary in order for \textsc{Scale-Free MD} to
have a sublinear regret. We show necessity of this assumption on two
counter-examples.  In these counter-examples, we consider strongly convex
regularizers $R$ such that $\Breg_R(u,v)$ is not bounded and we construct
sequences of loss vectors $\ell_1, \ell_2, \dots, \ell_T$ such that
$\|\ell_1\|_* = \|\ell_2\|_* = \dots = \|\ell_T\|_* = 1$ and \textsc{Scale-Free
MD} has regret $\Omega(T)$ or worse.

The first counter-example is stated as
Theorem~\ref{theorem:first-counter-example} below; proof is
in~\ref{section:mirror-descent-proofs}. The decision set is the whole space
$K=V$ and the regularizer is $R(w) = \frac{1}{2}\|w\|_2^2$. Note that $R(w)$ is
$1$-strongly convex with respect to $\|\cdot\|_2$ and the dual norm of
$\|\cdot\|_2$ is $\|\cdot\|_2$. The corresponding Bregman divergence is
$\Breg_R(u,v) = \frac{1}{2}\|u-v\|_2^2$. The counter-example constructs
sequence of unit-norm loss vectors in the one-dimensional subspace spanned by
the first vector of the standard orthnormal basis.  On such sequence,
\textsc{GIGA} and both versions of \textsc{AdaGrad MD} have predictions
identical to \textsc{Scale-Free MD}. Hence the lower bound applies to all four
algorithms.

@@

\begin{theorem}[Second Counter-Example]
\label{theorem:second-counter-example}
Let $d \ge 2$, let $V = \R^d$, and let $K = \{ w \in V ~:~ w_j \ge 0,
\sum_{j=1}^d w_j = 1 \}$ be the $d$-dimensional probability simplex.  For any
$T \ge 120$, there exists a sequence of loss vectors $\ell_1, \ell_2, \dots,
\ell_T \in V^*$ such that $\|\ell_1\|_\infty = \|\ell_2\|_\infty = \dots =
\|\ell_T\|_\infty = 1$ and \textsc{Scale-Free MD} with regularizer
$R(w) = \sum_{j=1}^d w_j \ln w_j$ satisfies
$$
\Regret_T \ge \frac{T}{6} \; .
$$
\end{theorem}
